---
title: "EDUC152, week 8 homework"
# author: "INSERT YOUR NAME HERE"
date: ""
urlcolor: blue
output: 
  #pdf_document:
    #pandoc_args: ["--extract-media", "."]
    #keep_md: yes
    #self_contained: true
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    self_contained: true
    #theme: default
    #highlight: tango
---

```{r setup, include=FALSE}

```

```{r, echo=FALSE, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

knitr::opts_chunk$set(collapse = TRUE, comment = "#>", highlight = TRUE, warning = FALSE, message = FALSE)
```

<span style="color:red"><b>Grade: /10<b></span>

# Overview  


The purpose of this short exercise is to give you some practice with the basics of multivariate regression. In this exercise you will:

- Run a regression model in *R* (code provided)
- Write out the population linear regression model
- Write out the OLS prediction line
  - with estimates
  - without estimates
- Interpret value of regression coefficients
- Calculate predicted values

# Load libraries and dataset
```{r}
# remove scientific notation
options(scipen=999)

##########
########## Libraries
##########

  library(tidyverse)
  library(labelled)
  library(haven)

##########
########## ELS:2002 data
##########

# RUN SCRIPT THAT CREATES STUDENT-LEVEL DATA FRAME CONTAINING ALL VARIABLES AND CREATES DATA FRAME WITH A SUBSET OF VARIABLES

  #NOTE: this script will take 30 seconds to a minute to run because loading a dataset w/ about 16,000 observations and 4,000 variables from a website

  source(file = url('https://github.com/anyone-can-cook/educ152/raw/main/scripts/els/read_els_by_pets.R'))
    #source(file = file.path('.','..','..','scripts','els','read_els_by_pets.R'))
      #list.files(path = file.path('.','..','..','scripts','els'))

# Create a dataframe df_els_stu_fac that has categorical variables as factor class variables rather than labelled class variables
  df_els_stu_fac <- as_factor(df_els_stu, only_labelled = TRUE) %>%
    # create a version of parent income that is in $1000s
    mutate(parent_income000 = parent_income/1000)
  # convert continuous variables we know we want numeric back to numeric
  for (v in c('bytxmstd','bytxrstd','f1txmstd','f3stloanamt','f3stloanpay','f3ern2011','f3tzrectrans','f3tzreqtrans','f3tzschtotal')) {
    df_els_stu_fac[[v]] <- df_els_stu[[v]]  
  }
```

# Run descriptive statistics and regression

Variables

- The dependent variable is high school reading test score, `bytxrstd`
  - variable label: `r attr(df_els_stu_fac$bytxrstd, which = 'label')`
- The continuous independent variable is `parent_income000`
  - variable label: `r attr(df_els_stu_fac$parent_income, which = 'label')`
  - *Note*. This is parent income in $thousands (e.g., value of 62.5 refers to \$62,500)
    - so a "one-unit" increase in this variable would be a \$1,000 increase in parent income
- The categorical independent variable is school control (e.g. public school, Catholic private school, or non-Catholic private school) `bysctrl`
  - variable label: `r attr(df_els_stu_fac$bysctrl, which = 'label')`

Your job in this section is just to run the provided code


- Descriptive statistics about variable in the model
```{r}
df_els_stu_fac %>% select(bytxrstd,parent_income000,bysctrl) %>% glimpse()

# dependent variable
df_els_stu_fac %>% summarize(
  mean_read_score = mean(bytxrstd, na.rm = TRUE),
  sd_read_score = sd(bytxrstd, na.rm = TRUE)
)

# continuous independent variable
df_els_stu_fac %>% summarize(
  mean_parent_income = mean(parent_income000, na.rm = TRUE),
  sd_parent_income = sd(parent_income000, na.rm = TRUE)
)

# categorical independent variable
df_els_stu_fac %>% count(bysctrl)

  # categorical independent variable, showing value of underlying integer values
 df_els_stu_fac %>% count(as.integer(bysctrl))
```

- Run regression model
```{r}
mod1 <- lm(formula = bytxrstd ~ parent_income000 + bysctrl, data = df_els_stu_fac %>% filter(f2enroll0506=='yes'))

summary(mod1)
```

# Questions for you to answer  

<span style="color:red"><b> /3<b></span>

#### 1. Write out the population linear regression  model  (make sure to define which variable (e.g., "parental income") is associated with which $X_{ki}$ in the model; and define unit of analysis if relevant).

- __YOUR ANSWER HERE__: 



    
<span style="color:red"><b> /2<b></span>

#### 2. Write out the OLS prediction line without estimate values and write out the OLS prediction line with estimate values.

__YOUR ANSWER HERE__: 




<span style="color:red"><b> /3<b></span> 

#### 3. Interpret the value of regression coefficients $\hat{\beta_1}$, $\hat{\beta_2}$, and $\hat{\beta_3}$ in words.

__YOUR ANSWER HERE__: 




<span style="color:red"><b> /1<b></span>

#### 4. Interpret the value of the regression coefficients $\hat{\beta_0}$ in words.

__YOUR ANSWER HERE__: 




<span style="color:red"><b> /1<b></span>  

#### 5. Calculate the predicted high school reading test score for a student who attended a non-Catholic private school and who has `parental_income000 = 150` (i.e., \$150,000); show your work  

__YOUR ANSWER HERE__: 




# Knit to html and submit exercise  

**Knit to html** by clicking the "Knit" button near the top of your RStudio window (icon with blue yarn ball) or drop down and select "Knit to HTML"

- Go to the [class website](https://anyone-can-cook.github.io/educ152/) and under the "Readings & Assignments" >> "Week 9" tab, click on the "Short exercise submission link"
- Submit both your html and .Rmd files 
- Use this naming convention "lastname_firstname_se" for your .Rmd (e.g. martin_patricia_se.Rmd)

