---
title: "EDUC152, Problem Set #1"
# author: "INSERT YOUR NAME HERE"
date: ""
urlcolor: blue
output: 
  #pdf_document:
    #pandoc_args: ["--extract-media", "."]
    #keep_md: yes
    #self_contained: true
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    self_contained: true
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", highlight = TRUE, warning = FALSE, message = FALSE)
```

### <span style="color:red"><b>Grade: /35<b></span>

# Overview

In this problem set, you will be working with the IPEDS dataset from lecture. You will be given the code to create several dataframes and plotting functions we will be working on in this problem set. Part I will give you practice working with distributions and sampling distributions and part II will give you practice with hypothesis testing. Both sections will include a mix of coding questions and conceptual questions where you will need to include a written response. Part III, will ask you to make a post on the class #problemsets channel about something you learned __or__ you can reply to another student's post.    

If you have any questions about the problem set, please also post them on the #problemsets slack channel. 


## Create data and load functions

Please run the code in the following chunk, which does the following:

- Loads libraries
- Loads and creates IPEDS data frame  (population)
- Creates data frame of generated variables (population)
- Creates sample versions of the IPEDS and gnerated data frames

_Note: code chunk omitted from PDF document using include = FALSE_
```{r, include= FALSE}
#-----------------------------------------------------------------------------
# Load libraries
#-----------------------------------------------------------------------------

  #install.packages('tidyverse') # if you haven't installed already
  #install.packages('labelled') # if you haven't installed already
  #install.packages('patchwork') # if you haven't installed already

library(tidyverse) # load tidyverse package
library(labelled) # load labelled package package
library(patchwork)

#-----------------------------------------------------------------------------
# Load ipeds dataset from course website url
#-----------------------------------------------------------------------------

load(file = url('https://github.com/anyone-can-cook/educ152/raw/main/data/ipeds/output_data/panel_data.RData'))

# Create ipeds data frame with fewer variables/observations
df_ipeds_pop <- panel_data %>%
  # keep data from fall 2019
  filter(year == 2019) %>%
  # which universities to keep:
    # 2015 carnegie classification: keep research universities (15,16,17) and master's universities (18,19,20)
  filter(c15basic %in% c(15,16,17,18,19,20)) %>%
  # which variables to keep
  select(instnm,unitid,opeid6,opeid,control,c15basic,stabbr,city,zip,locale,obereg, # basic institutional characteristics
         tuition6,fee6,tuition7,fee7, # avg tuition and fees for full-time grad, in-state and out-of-state
         isprof3,ispfee3,osprof3,ospfee3, # avg tuition and fees for MD, in-state and out-of-state
         isprof9,ispfee9,osprof9,ospfee9, # avg tuition and fees for Law, in-state and out-of-state
         chg4ay3,chg7ay3,chg8ay3) %>% # [undergraduate] books+supplies; off-campus (not with family) room and board; off-campus (not with family) other expenses
  # rename variables; syntax <new_name> = <old_name>
  rename(region = obereg, # revion
         tuit_grad_res = tuition6, fee_grad_res = fee6, tuit_grad_nres = tuition7, fee_grad_nres = fee7, # grad
         tuit_md_res = isprof3, fee_md_res = ispfee3, tuit_md_nres = osprof3, fee_md_nres = ospfee3, # md
         tuit_law_res = isprof9, fee_law_res = ispfee9, tuit_law_nres = osprof9, fee_law_nres = ospfee9, # law
         books_supplies = chg4ay3, roomboard_off = chg7ay3, oth_expense_off = chg8ay3) %>% # [undergraduate] expenses
  # create measures of tuition+fees
  mutate(
    tuitfee_grad_res = tuit_grad_res + fee_grad_res, # graduate, state resident
    tuitfee_grad_nres = tuit_grad_nres + fee_grad_nres, # graduate, non-resident
    tuitfee_md_res = tuit_md_res + fee_md_res, # MD, state resident
    tuitfee_md_nres = tuit_md_nres + fee_md_nres, # MD, non-resident
    tuitfee_law_res = tuit_law_res + fee_law_res, # Law, state resident
    tuitfee_law_nres = tuit_law_nres + fee_law_nres) %>% # Law, non-resident  
  # create measures of cost-of-attendance (COA) as the sum of tuition, fees, book, living expenses
  mutate(
    coa_grad_res = tuit_grad_res + fee_grad_res + books_supplies + roomboard_off + oth_expense_off, # graduate, state resident
    coa_grad_nres = tuit_grad_nres + fee_grad_nres + books_supplies + roomboard_off + oth_expense_off, # graduate, non-resident
    coa_md_res = tuit_md_res + fee_md_res + books_supplies + roomboard_off + oth_expense_off, # MD, state resident
    coa_md_nres = tuit_md_nres + fee_md_nres + books_supplies + roomboard_off + oth_expense_off, # MD, non-resident
    coa_law_res = tuit_law_res + fee_law_res + books_supplies + roomboard_off + oth_expense_off, # Law, state resident
    coa_law_nres = tuit_law_nres + fee_law_nres + books_supplies + roomboard_off + oth_expense_off) %>% # Law, non-resident    
  # keep only observations that have non-missing values for cost of off-campus room and board and cost of off-campus other expenses and non-missing values for tuit+fees
  filter(!is.na(roomboard_off) & !is.na(oth_expense_off) & !is.na(tuitfee_grad_nres))

# Add variable labels to the tuit+fees variables and coa variables
  # tuition + fees variables
    var_label(df_ipeds_pop[['tuitfee_grad_res']]) <- 'graduate, full-time, resident; avg tuition + required fees'
    var_label(df_ipeds_pop[['tuitfee_grad_nres']]) <- 'graduate, full-time, non-resident; avg tuition + required fees'
    var_label(df_ipeds_pop[['tuitfee_md_res']]) <- 'MD, full-time, state resident; avg tuition + required fees'
    var_label(df_ipeds_pop[['tuitfee_md_nres']]) <- 'MD, full-time, non-resident; avg tuition + required fees'
    var_label(df_ipeds_pop[['tuitfee_law_res']]) <- 'Law, full-time, state resident; avg tuition + required fees'
    var_label(df_ipeds_pop[['tuitfee_law_nres']]) <- 'Law, full-time, non-resident; avg tuition + required fees'
    
  # COA variables
    var_label(df_ipeds_pop[['coa_grad_res']]) <- 'graduate, full-time, state resident COA; == tuition + fees + (ug) books/supplies + (ug) off-campus room and board + (ug) off-campus other expenses'
    var_label(df_ipeds_pop[['coa_grad_nres']]) <- 'graduate, full-time, non-resident COA; == tuition + fees + (ug) books/supplies + (ug) off-campus room and board + (ug) off-campus other expenses'
    var_label(df_ipeds_pop[['coa_md_res']]) <- 'MD, full-time, state resident COA; == tuition + fees + (ug) books/supplies + (ug) off-campus room and board + (ug) off-campus other expenses'
    var_label(df_ipeds_pop[['coa_md_nres']]) <- 'MD, full-time, non-resident COA; == tuition + fees + (ug) books/supplies + (ug) off-campus room and board + (ug) off-campus other expenses'
    var_label(df_ipeds_pop[['coa_law_res']]) <- 'Law, full-time, state resident COA; == tuition + fees + (ug) books/supplies + (ug) off-campus room and board + (ug) off-campus other expenses'
    var_label(df_ipeds_pop[['coa_law_nres']]) <- 'Law, full-time, non-resident COA; == tuition + fees + (ug) books/supplies + (ug) off-campus room and board + (ug) off-campus other expenses'

df_ipeds_pop %>% glimpse()

#-----------------------------------------------------------------------------
# Create data frame of generated variables, with each variable meant to represent the entire population
#-----------------------------------------------------------------------------

num_obs <- 10000

# Generate normal distribution w/ custom mean and sd
set.seed(124)
norm_dist <- rnorm(n = num_obs, mean = 50, sd = 5)

# Generate right-skewed distribution
set.seed(124)
rskew_dist <- rbeta(n = num_obs, shape1 = 2, shape2 = 5)

# Generate left-skewed distribution
set.seed(124)
lskew_dist <- rbeta(n = num_obs, shape1 = 5, shape2 = 2)

# Generate standard normal distribution (default is mean = 0 and sd = 1)
set.seed(124)
stdnorm_dist <- rnorm(n = num_obs, mean = 0, sd = 1)  # equivalent to rnorm(10)

# Create dataframe
df_generated_pop <- data.frame(norm_dist, rskew_dist, lskew_dist, stdnorm_dist)

# drop individual objects associated with each variable
rm(norm_dist,rskew_dist,lskew_dist,stdnorm_dist)
rm(num_obs)

#-----------------------------------------------------------------------------
# Create sample versions of generated population data frame and IPEDS #-----------------------------------------------------------------------------

# create sample version of our generated data
  set.seed(124) # set seed so that everyone ends up with the same random sample
  
  df_generated_sample <- df_generated_pop %>% sample_n(size = 200)
  df_generated_sample %>% glimpse()


# create sample version of our ipeds data

  set.seed(124) # set seed so that everyone ends up with the same random sample
  
  df_ipeds_sample <- df_ipeds_pop %>% sample_n(size = 200) 


#-----------------------------------------------------------------------------
# Create function to generate plots of variable distributions
#-----------------------------------------------------------------------------

plot_distribution <- function(data_vec, plot_title = '') {
  p <- ggplot(as.data.frame(data_vec), aes(x = data_vec)) +
    ggtitle(plot_title) + xlab('') + ylab('') +
    geom_histogram(aes(y = ..density..), alpha = 0.4, position = 'identity') +
    geom_density() +
    geom_vline(aes(xintercept = mean(data_vec, na.rm = T), color = 'mean'),
               linetype = 'dotted', size = 0.8, alpha = 0.8) +
    geom_vline(aes(xintercept = median(data_vec, na.rm = T), color = 'median'),
               linetype = 'dotted', size = 0.8, alpha = 0.8) +
    scale_color_manual(name = 'Statistics',
                       labels = c(paste('Mean:', round(mean(data_vec, na.rm = T), 2),
                                        '\nStd Dev:', round(sd(data_vec, na.rm = T), 2)),
                                  paste('Median:', round(median(data_vec, na.rm = T), 2))),
                       values = c(mean = 'blue', median = 'red')) +
    theme(plot.title = element_text(size = 10, face = 'bold', hjust = 0.5),
          legend.title = element_text(size = 9, face = 'bold'),
          legend.text = element_text(size = 8))

  p
}

#-----------------------------------------------------------------------------
# Write function to get the sampling distribution from a variable (defaults equal 500 samples of size 200)
#-----------------------------------------------------------------------------

get_sampling_distribution <- function(data_vec, num_samples = 500, sample_size = 200) {
  sample_means <- vector(mode = 'numeric', num_samples)

  for (i in 1:length(sample_means)) {
    samp <- sample(data_vec, sample_size)
    sample_means[[i]] <- mean(samp, na.rm = T)
  }

  sample_means
}

#-----------------------------------------------------------------------------
# Write Function to generate sampling distribution (with t-test value) assuming null hypothesis is correct
#-----------------------------------------------------------------------------


plot_t_distribution <- function(data_vec, mu, alpha = 0.05, alternative = 'two.sided', plot_title = '', shade_rejection = T, shade_pval = T, stacked = F) {
  
  data_vec <- na.omit(data_vec)
  
  # Calculate t-statistics
  sample_size <- length(data_vec)
  deg_freedom <- sample_size - 1
  xbar <- mean(data_vec)
  s <- sd(data_vec)
  
  std_err <- s / sqrt(sample_size)
  t <- (xbar - mu) / std_err
  
  # Calculate critical value and p-value
  if (alternative == 'less') {  # left-tailed
    cv_lower <- qt(p = alpha, df = deg_freedom, lower.tail = T)
    cv_legend <- round(cv_lower, 2)
    cv_legend2 <- round(cv_lower * std_err + mu, 2)
    pval <- round(pt(q = t, df = deg_freedom, lower.tail = T), 4)
  } else if (alternative == 'greater') {  # right-tailed
    cv_upper <- qt(p = alpha, df = deg_freedom, lower.tail = F)
    cv_legend <- round(cv_upper, 2)
    cv_legend2 <- round(cv_upper * std_err + mu, 2)
    pval <- round(pt(q = t, df = deg_freedom, lower.tail = F), 4)
  } else {  # two-tailed
    cv_lower <- qt(p = alpha / 2, df = deg_freedom, lower.tail = T)
    cv_upper <- qt(p = alpha / 2, df = deg_freedom, lower.tail = F)
    cv_legend <- str_c('\u00B1', round(cv_upper, 2))
    cv_legend2 <- str_c(round(cv_lower * std_err + mu, 2), ' & ', round(cv_upper * std_err + mu, 2))
    pval_half <- round(pt(q = t, df = deg_freedom, lower.tail = t < 0), 4)
    pval <- str_c(pval_half, ' + ', pval_half, ' = ', 2 * pval_half)
  }
  
  # Plot t-distribution
  p <- ggplot(data.frame(x = -c(-4, 4)), aes(x)) +
    ggtitle(plot_title) + xlab('') + ylab('') +
    stat_function(fun = dt, args = list(df = deg_freedom), xlim = c(-4, 4))
  
  # Shade rejection region using critical value
  if (alternative != 'greater') {
    p <- p + geom_vline(aes(xintercept = cv_lower, color = 'cval'),
                        linetype = 'dotted', size = 0.8, alpha = 0.8)
    
    if (shade_rejection) {
      p <- p + stat_function(fun = dt, args = list(df = deg_freedom),
                             xlim = c(-4, cv_lower),
                             geom = 'area', alpha = 0.3, fill = 'red')
    }
    
    if (shade_pval) {
      p <- p + stat_function(fun = dt, args = list(df = deg_freedom),
                             xlim = c(-4, if_else(alternative == 'two.sided', -abs(t), t)),
                             geom = 'area', alpha = 0.3, fill = 'blue')
    }
  }
  if (alternative != 'less') {
    p <- p + geom_vline(aes(xintercept = cv_upper, color = 'cval'),
                        linetype = 'dotted', size = 0.8, alpha = 0.8)
    
    if (shade_rejection) {
      p <- p + stat_function(fun = dt, args = list(df = deg_freedom),
                             xlim = c(cv_upper, 4),
                             geom = 'area', alpha = 0.3, fill = 'red')
    }
    
    if (shade_pval) {
      p <- p + stat_function(fun = dt, args = list(df = deg_freedom),
                             xlim = c(if_else(alternative == 'two.sided', abs(t), t), 4),
                             geom = 'area', alpha = 0.3, fill = 'blue')
    }
  }
  
  # Legend text
  legend_text <- c('t-statistics / p-value', 'critical value / alpha')
  
  if (stacked) {
    legend_text <- c(str_c('t-statistics: ', round(t, 2),
                     '\n(p-value: ', str_extract(pval, '[\\d.-]+$'), ')'),
                     str_c('Critical value: ', cv_legend,
                     '\n(alpha: ', round(alpha, 2), ')'))
  }
  
  stats_text <- c(str_c('t-statistics: ', round(t, 2)),
                  str_c('SE: ', round(std_err, 2)),
                  str_c('p-value: ', pval),
                  str_c('Critical value: ', cv_legend),
                  str_c('alpha: ', round(alpha, 2)))
  
  if (!stacked) {
    p <- p +
      annotate('text', size = 9*5/14, x = 4.84, y = 0.14, hjust = 0,
               label = 'bold(Statistics)', parse = T) +
      annotate('text', size = 8*5/14, x = 4.89, y = 0:4 * -0.015 + 0.12, hjust = 0,
               label = stats_text)
  }
  
  # Label plot
  p <- p +
    geom_vline(aes(xintercept = t, color = 'tstat'),
               linetype = 'dotted', size = 0.8, alpha = 0.8) +
    scale_x_continuous(sec.axis = sec_axis(trans = ~ . * std_err + mu)) +
    scale_color_manual(name = if_else(stacked, 'Statistics', 'Legend'),
                       breaks = c('tstat', 'cval'),
                       labels = legend_text,
                       values = c(tstat = 'blue', cval = 'red')) +
    theme(plot.title = element_text(size = 10, face = 'bold', hjust = 0.5),
          plot.margin = unit(c(5.5, if_else(stacked, 5.5, 30), 5.5, 5.5), 'pt'),
          legend.title = element_text(size = 9, face = 'bold'),
          legend.text = element_text(size = 8)) +
    coord_cartesian(xlim = c(-4, 4),
                    clip = 'off')

  p
}
```



## Part I: Distributions and sampling distribution

<span style="color:red"><b>/5<b></span> 

#### 1. Use the function we created in the code chunk above `plot_distribution()` to plot the distribution of the variable `norm_dist` from the data frame `df_generated_pop`
```{r}

```

What is the standard deviation and interpret this value in words

- __YOUR ANSWER HERE__: 

Does the distribution above have a normal, left-skewed, or right-skwed shape? Why?

- __YOUR ANSWER HERE__: 

What is the "empirical rule"? Drawing from the empirical rule, what percentage of observations in the above distribution have values between 45 and 55? between 40 and 60? between 35 and 65?  

- Note: Make sure you answer all parts of the question.

- __YOUR ANSWER HERE__: 


<span style="color:red"><b>/2<b></span> 

#### 2. Use the function we created `plot_distribution()` to plot the distribution of the variable `tuitfee_grad_nres` from the data frame `df_ipeds_pop`. 

- Note: the data frame `df_ipeds_pop` contains data on the entire population of research/master's universities, whereas the data frame `df_ipeds_sample` contains data on a random sample of universities from that population.
```{r}

```

Does this variable appear to have a normal, left-skewed, or right-skewed distribution? why?

- __YOUR ANSWER HERE__:  


<span style="color:red"><b>/2<b></span>  

#### 3. Use the function we created `plot_distribution()` to plot the distribution of the variable `tuitfee_grad_nres` from the data frame `df_ipeds_sample`

- Note: the data frame `df_ipeds_pop` contains data on the entire population of research/master's universities, whereas the data frame `df_ipeds_sample` contains data on a random sample of universities from that population.
```{r}

```

Does this variable appear to have a normal, left-skewed, or right-skewed distribution? why?

- __YOUR ANSWER HERE__: 


<span style="color:red"><b>/2<b></span> 

#### 4. What is a sampling distribution? What is a sampling distribution of a sample mean?

- __YOUR ANSWER HERE__: 

<span style="color:red"><b>/6<b></span>   

#### 5. Run the following code, which does the following:

- Takes 1000 random samples of sample size n=200 from the data frame `df_ipeds_pop`.
- For each random sample, calculates the sample mean of variable `tuitfee_grad_nres`.
- Plots the sampling distribution of the sample mean of variable  `tuitfee_grad_nres`.

```{r}
set.seed(124)
get_sampling_distribution(data_vec = df_ipeds_pop$tuitfee_grad_nres, num_samples = 1000, sample_size = 200) %>%
  plot_distribution(plot_title = "Sampling Distribution of the Sample mean of out-of-state graduate tuition and fees")

#same as above
#plot_distribution(get_sampling_distribution(data_vec = df_ipeds_pop$tuitfee_grad_nres, num_samples = 1000, sample_size = 200),plot_title = "sampling distribution of sample mean of tuitfee_grad_nres")
```

Answer the following questions with respect to the above plot (one sentence or less for each answer):

- What does each observation in the above plot represent?
  - __YOUR ANSWER HERE__: 
  
- Would you describe the shape of the above distribtuion as (approximately) normal, left-skwed, or right-skewed?
  - __YOUR ANSWER HERE__:
  
- Define what the concept "standard error" mean (referrring to sampling distribution of sample mean)?
  - __YOUR ANSWER HERE__: 
  
- Why are the concepts "standard error" and "standard deviation of the sampling distribution" equivalent?
  - __YOUR ANSWER HERE__: 
  
- Interpret the value of standard error in the above plot in words
  - __YOUR ANSWER HERE__: 
  
- Write the formula for sample standard error and state what each component of the formula refers to (e.g., `n` refers to sample size)
  - __YOUR ANSWER HERE__: 
  

<span style="color:red"><b>/2<b></span>    

#### 6. Run the following code, which does the following:

- Takes 1000 random samples of sample size n=20 from the data frame `df_ipeds_pop`
- For each random sample, calculates the sample mean of variable `tuitfee_grad_nres`
- Plots the sampling distribution of the sample mean of variable  `tuitfee_grad_nres`

```{r}
set.seed(124)
get_sampling_distribution(data_vec = df_ipeds_pop$tuitfee_grad_nres, num_samples = 1000, sample_size = 20) %>%
  plot_distribution(plot_title = "Sampling distribution of sample mean of tuitfee_grad_nres")
#,plot_title = 'Sampling distribution')
```

Answer the following questions with respect to the above plot (one sentence or less for each answer):

- Interpret the value of standard error in words
  - __YOUR ANSWER HERE__: 
  
- Why is the standard error from this sampling distribution (each sample has sample size n=20) larger than the sampling distribution from the previous example (each sample has sample size n=200)?  
  - __YOUR ANSWER HERE__:


<span style="color:red"><b>/2<b></span>   

#### 7. Run the following code, which does the following:

- Plots the population distribution of the variable `tuitfee_grad_nres`
- Plots the distribution of the variable  `tuitfee_grad_nres` from one sample
- Plots the sampling distribution of the sample mean for the variable `tuitfee_grad_nres`
```{r}
set.seed(124)
plot_distribution(df_ipeds_pop$tuitfee_grad_nres, plot_title = 'Population distribution') +
  plot_distribution(df_ipeds_sample$tuitfee_grad_nres, plot_title = 'Single sample distribution') +
  plot_distribution(get_sampling_distribution(data_vec = df_ipeds_pop$tuitfee_grad_nres, num_samples = 1000, sample_size = 200),plot_title = "sampling distribution of sample mean of tuitfee_grad_nres") +
  plot_layout(ncol = 1)
```

State the central limit theorem in your own words and explain why it is important for hypothesis testing

- __YOUR ANSWER HERE__: 


## Part II: Hypothesis testing

In this section we will be testing a hypothesis about the variable off-campus room and board (`roomboard_off`).

Here is how IPEDS defines concepts related to room and board and other expenses, frome the IPEDS "Student Charges for Full Academic Year" 2019-20 academic year data dictionary [[LINK]](https://nces.ed.gov/ipeds/datacenter/data/IC2019_AY_Dict.zip):

- "Room charges"
  - The charges for an academic year for rooming accommodations for a typical student sharing a room with one other student. 

- "Board charges"
  - The charge for an academic year for meals, for a specified number of meals per week. 
- "Other expenses"
  - The amount of money (estimated by the financial aid office) needed by a student to cover expenses such as laundry, transportation, entertainment, and furnishings. (For the purpose of this survey room and board and tuition and fees are not included.) 
- _Note that most of these variables seem to be defined for an academic year rather than a 12-month calendar year._


Here, We have included some code to help you get to know the data. Just run this code and take a look at the output

Print observations for UC campuses
```{r}
df_ipeds_pop %>%
  # keep UC campuses
  filter(unitid %in% c(110398,110635,110644,110653,110662,110671,110680,110699,110705,110714,445188,110699,110398)) %>%
  select(instnm,city,locale,roomboard_off,oth_expense_off) %>% as_factor()
```

The variable `locale` categorizes universities by city/suburb/town/rural and by city size
```{r}
#df_ipeds_pop %>% count(locale)
df_ipeds_pop %>% count(locale) %>% as_factor()
```

Average cost of off-campus room & board
```{r}
mean(df_ipeds_pop$roomboard_off, na.rm = TRUE)

#alternative approach for calculating mean room and board
df_ipeds_pop %>% summarize(mean_roomboard_off = mean(roomboard_off, na.rm = TRUE))
```

Average cost of off-campus room & board, separately for each value of `locale`
```{r}
df_ipeds_pop %>% group_by(locale) %>% #creates a separate group for each locale
  summarize(
    sample_size = n(), #gets the count of colleges in each locale
    mean_roomboard_off = mean(roomboard_off, na.rm = TRUE) #calculate the mean room and board costs for each locale
    ) %>% as_factor() #return as factor
```


<span style="color:red"><b>/5<b></span>   

#### 1. What are the five steps in hypothesis testing? for each step, provide a one-sentence description.

- __YOUR ANSWER HERE__:
 

<span style="color:red"><b>/2<b></span>  

#### 2. Hypothesis testing steps

In the below questions, you will conduct hypothesis testing steps to answer the research question, "Is the population mean of off-campus room & board equal to $10,000?" You will be using the variable `roomboard_off` from the data frame `df_ipeds_sample`, which is a single random sample from the population data frame `df_ipeds_pop`. You will use a two-sided alternative hypothesis with an alpha level (rejection region) of `.05`.

- State the null and alternative (two-sided) hypothesis
  - __YOUR ANSWER HERE__:
  
  
<span style="color:red"><b>/1<b></span> 

#### 3. Use the `t.test()` function to calculate the test statistic


```{r}

```


<span style="color:red"><b>/4<b></span>  

#### 4. Use function `plot_t_distribution()` we created above to plot the sampling distribution under the assumption that $H_0$ is true.
```{r}

```

- Interpret the t-value in words and interpret the p-value in words.  
  - __YOUR ANSWER HERE__: 

- State the conclusion about your hypothesis test.  
  - __YOUR ANSWER HERE__: 
  

## Part III: Post a comment/question

<span style="color:red"><b>/2<b></span> 

- Go to the class [#problemsets ](https://app.slack.com/client/T01S30RQY2E/C01SH5DSNEN) channel and create a new post.
- You can either:
  - Share something you _learned_ or a _question_ from this problem set. Make sure to mention the instructors (@ozanj, @Patricia Martín).
  - Respond to a post made by another student.


# Knit to html and submit problem set  

**Knit to html** by clicking the "Knit" button near the top of your RStudio window (icon with blue yarn ball) or drop down and select "Knit to HTML"

- Go to the [class website](https://anyone-can-cook.github.io/educ152/) and under the "Readings & Assignments" >> "Week 3" tab, click on the "Problem set 1 submission link"
- Submit both your html and .Rmd files 
- Use this naming convention "lastname_firstname_ps#" for your .Rmd (e.g. martin_patricia_ps1.Rmd)
